---
title: "Sales Forecasts"
author: "Yun Xing"
date: '2023-05-23'
output: html_document
---


#### Superstore Sales Dataset - https://www.kaggle.com/datasets/rohitsahoo/sales-forecasting


```{r cars}
suppressMessages(library(tidyverse))
suppressMessages(library(maps))
suppressMessages(library(lubridate))
suppressMessages(library(viridis))
suppressMessages(library(janitor))
suppressMessages(library(scales))
suppressMessages(library(corrplot))
suppressMessages(library(RColorBrewer))
suppressMessages(library(wesanderson))
suppressMessages(library(htmltools))
suppressMessages(library(webr))
suppressMessages(library(fastDummies))
```


## Data Cleaning


```{r}
sales<-read.csv("train.csv")
head(sales,3)
```

```{r}
dim(sales)
```


```{r}
summary(sales)
```
```{r}
str(sales)
```

```{r}
colnames(sales)
```


#### check duplicates and missing values 


```{r}
sum(duplicated(sales))
```

```{r}
colSums(is.na(sales))
```

```{r}
sales[is.na(sales$Postal.Code), ]
```

#### all in vermont. so we fill in the postal code 05041

```{r}
sales$Postal.Code<-replace_na(sales$Postal.Code,05401)
```

```{r}
colSums(is.na(sales))

```

#### modify dates

```{r}

sales$Order.Date<-as.Date(sales$Order.Date,'%d/%m/%Y')
sales$Ship.Date<-as.Date(sales$Ship.Date, '%d/%m/%Y')
class(sales$Order.Date)
class(sales$Ship.Date)

```


## EDA


```{r}
cat("Number of unique Order IDs:\n")
length(unique(sales$Order.ID))

cat("Unique Ship Mode:\n")
unique(sales$Ship.Mode)

cat("Number of unique Customer IDs:\n")
length(unique(sales$Customer.ID))

cat("Number of unique Customer name:\n")
length(unique(sales$Customer.Name))

cat("Unique segment:\n")
unique(sales$Segment)

cat("Unique country:\n")
unique(sales$Country)

cat("Number of unique Cities:\n")
length(unique(sales$City))

cat("Unique Region:\n")
unique(sales$Region)
```
```{r}
#Top 5 customers by value
### assuming no customers have same name because length of unique customer.id is same as length of unique customer.name

customer_sales <- sales %>% 
    group_by(Customer.Name) %>% 
  summarise(sales=round(sum(Sales),2)) %>% 
  arrange(desc(sales))

head(customer_sales,5)

```

```{r}
ggplot(head(customer_sales,10), aes(x=fct_reorder(Customer.Name,sales), y=sales)) +geom_col(width = 0.7, fill='skyblue', col="black") +  labs(title = " Top 10 Most Valuable Customers", y="Sales", x="Customer") + 
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 20))+
geom_text(aes(label = sales), size = 3.5, vjust = -0.5, hjust = 0.5, angle = 0)


```


```{r}
# mean sales amount
mean(sales$Sales)
```

```{r}
#sales by order.id
sales_by_order<-sales %>% 
  group_by(Order.ID) %>% 
  summarise(sales=round(sum(Sales),2)) %>% 
  arrange(desc(sales))
```
```{r}
# mean sales amount by order id 
mean(sales_by_order$sales)
```

```{r}
# sales distribution
ggplot(sales_by_order, aes(x=sales)) + 
geom_histogram(aes(y=..density..),fill='white', color='black')+
geom_density(fill='skyblue', alpha=0.4,size=0.6) + scale_x_log10() + 
  labs(title = "Sales Distribution")
```

```{r}
# sales date analysis
sales<-sales %>% 
  mutate(order_year=format(Order.Date,"%Y"),.after=Order.Date)

sales<-sales %>% 
  mutate(order_month=format(Order.Date,"%b"),.after=order_year)

sales<-sales %>% 
  mutate(order_day=format(Order.Date,"%a"),.after=order_month)

#Converting month and day to factors for later use
sales$order_month<-factor(sales$order_month,
                          levels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
sales$order_day<-factor(sales$order_day,
                        levels = c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"))
```

```{r}
head(sales,3)
```


```{r}
sales_by_day<-sales %>% 
    group_by(order_day) %>% 
  summarise(sales=round(sum(Sales),0)) 

ggplot(sales_by_day, aes(x=order_day,y=sales)) +
geom_col(fill="skyblue", alpha=0.7, width = 0.5, color="black") + 
labs(x="",y="Sales", title="Sales by Day") + 
theme() + theme(plot.title = element_text(hjust=0.5)) + 
scale_y_continuous(labels = comma) +
  geom_text(aes(label = sales), size = 3.5, vjust = -0.5, hjust = 0.5, angle = 0)
```
```{r}
sales_by_month<-sales %>% 
  group_by(order_month) %>% 
  summarise(sales=round(sum(Sales),0))

ggplot(sales_by_month, aes(x=order_month,y=sales)) +
geom_col(fill="skyblue", alpha=0.7, width = 0.7, color="black") + 
labs(x="",y="Sales", title="Sales by Month") + 
theme(plot.title = element_text(hjust=0.5)) +
scale_y_continuous(labels = comma)+
  geom_text(aes(label = sales), size = 3.5, vjust = -0.5, hjust = 0.5, angle = 0)

```

```{r}
unique(sales$order_year)
sales$order_year<-factor(sales$order_year,levels = c(2015,2016,2017,2018))
```
```{r}
sales_by_year<-sales %>% 
  group_by(order_year) %>% 
  summarise(sales=round(sum(Sales),0))

ggplot(sales_by_year, aes(x = order_year, y = sales)) +
  geom_col(fill = "skyblue", alpha = 0.7, width = 0.5, color = "black") +
  labs(x = "", y = "Sales", title = "Sales by Month") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma) +
  geom_text(aes(label = sales), size = 3.5, vjust = -0.5, hjust = 0.5, angle = 0)

```

```{r}

ship_mode_counts <- table(sales$Ship.Mode)

pie(ship_mode_counts, labels = paste0(names(ship_mode_counts), ": ", round(prop.table(ship_mode_counts) * 100,0), "%"))

```



```{r}
sales_by_seg <- sales %>%
  group_by(Segment) %>%
  summarise(sales = sum(Sales))
```
```{r}
ggplot(sales_by_seg, aes(x = Segment, y = sales)) +
  geom_col(fill = 'skyblue', alpha = 0.8,width = 0.5) +
  labs(x = "Segment", y = "Sales") +
  #theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 1)) +
  geom_text(aes(label = sales), size = 3.5, vjust = -0.5, hjust = 0.5, angle = 0)
  coord_flip()
```
```{r}
pie(sales_by_seg$sales, labels = paste0(sales_by_seg$Segment, ": ", round(prop.table(sales_by_seg$sales) * 100, 0), "%"))

```

```{r}
suppressMessages(library(ggthemes))

foo <- sales %>%
  group_by(State, order_month, order_year) %>% 
  summarise(sales = sum(Sales))

top_states <- foo %>%
  group_by(State) %>%
  summarise(total_sales = sum(sales)) %>%
  top_n(3, total_sales) %>%
  pull(State)

filtered_data <- foo %>%
  filter(State %in% top_states)

ggplot(filtered_data, aes(x = paste(order_month, order_year, sep = "-"), y = sales, group = State, color = State)) +
  geom_line() +
  theme_tufte() +
  labs(x = "Year", y = "Sales", color = "State") +
  scale_x_discrete(labels = function(x) gsub("-", " ", x)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


```


```{r}
top10_states <- foo %>%
  group_by(State) %>%
  summarise(total_sales = sum(sales)) %>%
  top_n(10, total_sales) %>%
  arrange(total_sales)

ggplot(top10_states, aes(x = reorder(State, desc(total_sales)), y = total_sales)) +
  geom_bar(stat = "identity", alpha = 0.7, fill = "pink") +
  geom_text(aes(label = total_sales), vjust = -0.5, hjust = 1, size = 3, color = "black") +
  labs(x = "State", y = "Total Sales") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()+
  theme_minimal()

```

```{r}
sales_by_region<-sales %>% 
  group_by(Region) %>% 
  summarise(sales=round(sum(Sales),0))

PieDonut(sales_by_region,aes(Region,count=sales),
         title = "Sales by Region",r0=0.6) +
        scale_fill_manual(values = c("skyblue", "grey", "pink", "coral", "purple"))
```

```{r}
sales_by_subcat<-sales %>% 
  group_by(Category,Sub.Category) %>% 
  summarise(sale=sum(Sales))

PieDonut(sales_by_subcat, 
         aes(Category,Sub.Category,count=sale),
         title = "Sales by Category and Sub-Category")


```

#### Correlations analysis


```{r}

cat<-sales[,c("Category","Sales")]

catdummy<-dummy_cols(cat)
catdummy<-catdummy %>% 
  rename(furniture=3, office_supplies=4, technology=5)
catdummy<-catdummy[,2:5]
#convert columns data type to numeric
catdummy$furniture<-as.numeric(catdummy$furniture)
catdummy$office_supplies<-as.numeric(catdummy$office_supplies)
catdummy$technology<-as.numeric(catdummy$technology)
#create a correlation matrix
corcatdummy<-cor(catdummy)
#plotting correlation matrix
corrplot(corcatdummy, type = "upper", method = "color",
         addCoef.col = "black",
         tl.col = "black", diag = FALSE, order = "hclust")


```
```{r}
seg<-sales[,c("Segment","Sales")]

segdummy<-dummy_cols(seg)
head(segdummy)
segdummy<-segdummy %>% 
  rename(consumer=3,corporate=4, home_office=5)

segdummy$consumer<-as.numeric(segdummy$consumer)
segdummy$corporate<-as.numeric(segdummy$corporate)
segdummy$home_office<-as.numeric(segdummy$home_office)
segdummy<-segdummy[,2:5]

corsegdummy<-cor(segdummy)

corrplot(corsegdummy, type = "upper", method = "color",
         addCoef.col = "black",
         tl.col = "black", diag = FALSE, order = "hclust")
```

```{r}
ship<-sales[,c("Ship.Mode","Sales")]

shipdummy<-dummy_cols(ship)
shipdummy<-shipdummy[,2:6] %>% 
  rename(first_class=2, same_day=3, second_class=4)

shipdummy$first_class<-as.numeric(shipdummy$first_class)
shipdummy$same_day<-as.numeric(shipdummy$same_day)
shipdummy$second_class<-as.numeric(shipdummy$second_class)

corshipdummy<-cor(shipdummy)

corrplot(corshipdummy, type = "upper", method = "color",
         addCoef.col = "black",
         tl.col = "black", diag = FALSE, order = "hclust")
```


```{r}
re<-sales[,c("Region","Sales")]

redummy<-dummy_cols(re)
redummy<-redummy[,2:6] %>% 
  mutate_all(as.numeric)

corredummy<-cor(redummy)

corrplot(corredummy, type = "upper", method = "color",
         addCoef.col = "black",
         tl.col = "black", diag = FALSE, order = "hclust")
```

#### Date plots 

```{r}
sales_small <- sales[c("Order.Date", "Sales")]

hist(sales_small$Order.Date, breaks = 18,
     main = "Order Date Histogram",
     xlab = "Order Date",
     ylab = "Count")


```

## Predict sales for most valuable states (Top 3) in 2018.

```{r}
foo <- sales %>%
  group_by(State, Order.Date, order_month, order_year) %>% 
  summarise(sales = sum(Sales))

top_states <- foo %>%
  group_by(State) %>%
  summarise(total_sales = sum(sales)) %>%
  top_n(3, total_sales) %>%
  pull(State)

df <- foo %>%
  filter(State %in% top_states)

```


```{r}
dim(df)
```


```{r}
df$Order.Date <- floor_date(df$Order.Date, "month")
```
```{r}
df<- df %>%
  group_by(State, Order.Date) %>% 
  summarise(sales = sum(sales))
```
```{r}
ca_df <- subset(df, State == "California", select = c(Order.Date, sales))
ny_df <- subset(df, State == "New York", select = c(Order.Date, sales))
tx_df <- subset(df, State == "Texas", select = c(Order.Date, sales))
```

```{r}
library(xts)

ca_xts <- as.xts(ca_df$sales, order.by = ca_df$Order.Date)
ny_xts <- as.xts(ny_df$sales, order.by = ny_df$Order.Date)
tx_xts <- as.xts(tx_df$sales, order.by = tx_df$Order.Date)
```


```{r}
plot(ca_xts, main = "CA Sales", ylab = "sales")
plot(ny_xts, main = "NY Sales", ylab = "sales")
plot(tx_xts, main = "TX Sales", ylab = "sales")
```

```{r}
ca_train <- ca_xts["/2017-12-01"]
ca_test <- ca_xts["2018-01-01/"]

ny_train <- ny_xts["/2017-12-01"]
ny_test <- ny_xts["2018-01-01/"]

tx_train <- tx_xts["/2017-12-01"]
tx_test <- tx_xts["2018-01-01/"]
```

```{r}
acf(ca_train)
acf(ny_train)
acf(tx_train)
```
```{r}
suppressMessages(library(tseries))
adf.test(ca_train)
kpss.test(ca_train)

```
```{r}
adf.test(ny_train)
kpss.test(ny_train)
```

```{r}
adf.test(tx_train)
kpss.test(tx_train)
```

### 1. Use forecast::auto.arima with no seasonality to get a baseline model 

```{r}
suppressMessages(library(forecast))

ca_base <- auto.arima(ca_train, seasonal = FALSE)
ca_base
```

```{r}
ny_base <- auto.arima(ny_train, seasonal = FALSE)
ny_base
```

```{r}
tx_base <- auto.arima(tx_train, seasonal = FALSE)
tx_base
```


CA
```{r}
a <- length(ca_test)
forecast_ca_base <- forecast(ca_base, h = a)
plot(forecast_ca_base, main = "Forecasts of CA monthly sales 2018 - baseline model")

residuals_ca <- ca_test - forecast_ca_base$mean
checkresiduals(residuals_ca)
```
```{r}
acc_ca_base <- accuracy(forecast_ca_base, ca_test)
acc_ca_base
```

NY

```{r}
forecast_ny_base <- forecast(ny_base, h = 12)
plot(forecast_ny_base, main = "Forecasts of NY monthly sales 2018 - baseline model")
acc_ny_base <- accuracy(forecast_ny_base, ny_test)
acc_ny_base

residuals_ny <- ny_test - forecast_ny_base$mean
checkresiduals(residuals_ny)
```

TX

```{r}
forecast_tx_base <- forecast(tx_base, h = 12)
plot(forecast_tx_base, main = "Forecasts of TX monthly sales 2018 - baseline model")
acc_tx_base <- accuracy(forecast_tx_base, tx_test)
acc_tx_base

residuals_tx <- tx_test - forecast_tx_base$mean
checkresiduals(residuals_tx)
```

#### predictions for these three are pretty bad with high RMSE. 

### 2. Seasonal ARIMA (SARIMA) - here seasonality is monthly

CA

```{r}
ca_SA <- auto.arima(ts(ca_train, frequency=12), seasonal = TRUE)
summary(ca_SA)

forecast_ca_SA <- forecast(ca_SA, h = 12)
acc_ca_SA <- accuracy(forecast_ca_SA, ca_test)
acc_ca_SA

plot(forecast_ca_SA, main = "Forecasts of CA monthly sales 2018 - SARIMA model")

residuals_ca1 <- ca_test - forecast_ca_SA$mean
checkresiduals(residuals_ca1)

```

NY

```{r}
ny_SA <- auto.arima(ts(ny_train, frequency=12), seasonal = TRUE)
summary(ny_SA)

forecast_ny_SA <- forecast(ny_SA, h = 12)
acc_ny_SA <- accuracy(forecast_ny_SA, ny_test)
acc_ny_SA

plot(forecast_ny_SA, main = "Forecasts of NY monthly sales 2018 - SARIMA model")

residuals_ny1 <- ny_test - forecast_ny_SA$mean
checkresiduals(residuals_ny1)
```
TX

```{r}
tx_SA <- auto.arima(ts(tx_train, frequency=12), seasonal = TRUE)
summary(tx_SA)

forecast_tx_SA <- forecast(tx_SA, h = 12)
acc_tx_SA <- accuracy(forecast_tx_SA, tx_test)
acc_tx_SA

plot(forecast_tx_SA, main = "Forecasts of TX monthly sales 2018 - SARIMA model")

residuals_tx1 <- tx_test - forecast_tx_SA$mean
checkresiduals(residuals_tx1)
```

### 3. Holt Winter 


CA

```{r}
HW.ca <- HoltWinters(ts(ca_train, frequency=12))
HW.ca

forecast_HW.ca <- forecast(HW.ca, h = 12)
acc_HW.ca <- accuracy(forecast_HW.ca, ca_test)
acc_HW.ca

plot(forecast_HW.ca, main = "Forecasts of CA monthly sales 2018 - HW model")

residuals_ca2 <- ca_test - forecast_HW.ca$mean
checkresiduals(residuals_ca2)
```

NY

```{r}
HW.ny <- HoltWinters(ts(ny_train, frequency=12))
HW.ny

forecast_HW.ny <- forecast(HW.ny, h = 12)
acc_HW.ny <- accuracy(forecast_HW.ny, ny_test)
acc_HW.ny

plot(forecast_HW.ny, main = "Forecasts of NY monthly sales 2018 - HW model")

residuals_ny2 <- ny_test - forecast_HW.ny$mean
checkresiduals(residuals_ny2)

```

TX

```{r}
HW.tx <- HoltWinters(ts(tx_train, frequency=12))
HW.tx

forecast_HW.tx <- forecast(HW.tx, h = 12)
acc_HW.tx <- accuracy(forecast_HW.tx, tx_test)
acc_HW.tx

plot(forecast_HW.tx, main = "Forecasts of TX monthly sales 2018 - HW model")

residuals_tx2 <- tx_test - forecast_HW.tx$mean
checkresiduals(residuals_tx2)
```

#### 4. arfima 
CA
```{r}
m41 <- forecast::arfima(ts(ca_train))

summary(m41)

pred_41 <- forecast(m41, h = length(ca_test))
resid_41 <- ca_test - pred_41$mean
checkresiduals(resid_41)
mean(resid_41)
```


NY 

```{r}
m42 <- forecast::arfima(ts(ny_train))

summary(m42)

pred_42 <- forecast(m42, h = length(ny_test))
resid_42 <- ny_test - pred_42$mean
checkresiduals(resid_42)
mean(resid_42)
```


TX
```{r}
m43 <- forecast::arfima(ts(tx_train))

summary(m43)

pred_43 <- forecast(m43, h = length(tx_test))
resid_43 <- tx_test - pred_43$mean
checkresiduals(resid_43)
mean(resid_43)
```


## Predict overall sales (all states aggregated)


```{r}
agg_sales <- sales %>%
  group_by(Order.Date) %>% 
  summarise(sales = sum(Sales))
```

```{r}
drange <- seq(as.Date("2015-01-03"), as.Date("2018-12-30"), by="day")

dall <- data.frame(Order.Date = drange)

dall <- merge(agg_sales, dall, by=c("Order.Date"), all.y=TRUE)

```
```{r}
dim(dall[is.na(dall$sales),])
```
```{r}
dim(dall)
```
```{r}
sales_xts <- as.xts(dall$sales, order.by = dall$Order.Date)
```


#### there are missing dates in original dataset. so after merging the dataset with the full date rage, there are missing values. now we fill them with 0. (because no sales occurred that day) 

```{r}

sales_xts <- na.fill(sales_xts, 0)

```
```{r}
sum(is.na(sales_xts))
```
```{r}
plot(sales_xts, main = "Total Sales by Date", ylab = "sales")
```
```{r}
train <- sales_xts["/2018-11-30"]
test <- sales_xts["2018-12-01/"]
```
```{r}
dim(train)
dim(test)
```

```{r}
acf(train)
```
```{r}
adf.test(train)
kpss.test(train)
```

### 1. forecast::auto.arima 


```{r}
base <- auto.arima(train, seasonal = FALSE)
base

a <- length(test)
forecast_base <- forecast(base, h = a)
plot(forecast_base, main = "Forecasts of Daily sales 2018 - baseline model")
```

### 2. seasonal arima 


```{r}
model.1 <- auto.arima(train, seasonal = TRUE)
summary(model.1)
```

```{r}
m1.forecast <- forecast(model.1, h=a)

plot(m1.forecast, main = "Forecasts of daily sales 2018 - sarima")
resid_m1 <- residuals(model.1)

mean(resid_m1) 
```

```{r}
checkresiduals(resid_m1)
```

### 3. holt winters
  
```{r}
hw <-  HoltWinters(train, gamma = FALSE, seasonal = 'additive')
hw
hw_forecast <-  forecast(hw, h = length(test))
resid_hw <- test - hw_forecast$mean

plot(hw_forecast, main = "Forecasts of daily sales 2018 - HoltWinters")

plot(hw_forecast$mean, main = "Forecasts of daily sales 2018(zoom in) - HoltWinters")

checkresiduals(resid_hw)
```
### 4. ARFIMA

```{r}

m4 <- forecast::arfima(ts(train))
AIC(m4)
summary(m4)
```

```{r}
pred_ARFIMA <- forecast(m4, h = length(test))

resid_arf <- test - pred_ARFIMA$mean

checkresiduals(resid_arf)

mean(resid_arf)

plot(pred_ARFIMA, main = "Forecasts of daily sales 2018 - ARFIMA")

plot(pred_ARFIMA$mean, main = "Forecasts of daily sales 2018(zoom in) - ARFIMA")
```

 
 